---
title: Validating Joint Analysis of Single-Cell Transcriptome and Epigenome Data Using
  scNMT-seq and SNARE-seq
author: "Joshua Welch"
date: "March 9, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SNARE-seq
We first read the data (gene body/promoter accessibility for ATAC-seq, gene expression for RNA-seq):
```{r read}
library(Matrix)
gene_body_atac = readRDS("GSE126074_AdBrainCortex_SNAREseq_chromatin_genes.counts.RDS")
rna = readMM("GSE126074_AdBrainCortex_SNAREseq_cDNA.counts.mtx.gz")
rn = read.table("GSE126074_AdBrainCortex_SNAREseq_cDNA.barcodes.tsv.gz",header=F)
rn = rn[,1]
cn = read.table("GSE126074_AdBrainCortex_SNAREseq_cDNA.genes.tsv.gz",header=F)
cn = cn[,1]
colnames(rna)=paste0("rna_",rn)
rownames(rna)=cn
```
We next normalize by total counts per cell, select variable genes (using RNA only), and perform iNMF:
```{r liger}
library(liger)
snare_liger = createLiger(list(rna=rna,atac=gene_body_atac))
snare_liger = normalize(snare_liger)
snare_liger = selectGenes(snare_liger,var.thresh = 0.075,datasets.use = 1)
snare_liger = scaleNotCenter(snare_liger)
snare_liger = optimizeALS(snare_liger,k=30,max.iters = 30,thresh=1e-6)
```
Now we perform quantile normalization, t-SNE, and joint clustering.
```{r align}
snare_liger = quantile_norm(snare_liger,ref_dataset = "rna",k=100)
snare_liger = runTSNE(snare_liger)
snare_liger = louvainCluster(snare_liger,resolution=1,k=100)
names(snare_liger@clusters)=rownames(snare_liger@H.norm)
```
Plotting indicates strong dataset alignment and confirms that gene body/promoter accessibility and gene expression are correlated.
```{r plotting,echo=FALSE}
plotByDatasetAndCluster(snare_liger)
#Glutamatergic, IT
plotGene(snare_liger,"Slc30a3")
#Glutamatergic, non-IT 
plotGene(snare_liger,"Foxp2")
#GABAergic
plotGene(snare_liger,"Gad2")
#OPC
plotGene(snare_liger,"Tnr")

```

